The \dealii{} Library, Version 9.3

### 2021-06-18 9.3/paper.tex

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Marc Fehling,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.3, 2021},
}

@@

\author[4]{Timo~Heister}
 \affil[4]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{tcleven/heister@clemson.edu}}}

\author[5]{Luca~Heltai}
\affil[5]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{luca.heltai@sissa.it}}}

\author[6]{Uwe~K{\"o}cher}
\affil[6]{Chair of Numerical Mathematics,
  Helmut-Schmidt-University,
  University of the Federal Armed Forces Hamburg,
  Holstenhofweg~85, 22043 Hamburg, Germany.
  {\texttt{uwe.koecher@hsu-hh.de}}}

 \author[7,8]{Martin~Kronbichler}
 \affil[7]{Institute for Computational Mechanics,
   Technical University of Munich,
   Boltzmannstr.~15, 85748 Garching, Germany.
   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
 \affil[8]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler@it.uu.se}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[7,10]{Peter Munch}
 \affil[10]{Institute of Materials Research, Materials Mechanics,
 Helmholtz-Zentrum Geesthacht,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hzg.de}}}

\author[11]{Jean-Paul~Pelteret}
\affil[11]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

@@

@@

\dealii{} version 9.3.0 was released May ??, 2021.
\todo{Update date}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.3. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
These major changes are discussed in detail in Section~\ref{sec:major}. There
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item Each non-artificial cell now has a globally unique index that can be queried
  for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
  via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
  can be used to efficiently index into global vectors storing data
  for each cell, rather than for each degree of freedom corresponding to a finite
  element field.

  \item Previously, functions and classes were marked using the 
  macro \texttt{DEAL\_II\_DEPRECATED} and then typically removed in
  the release after the one in which these deprecation notices were
  available to users. We have now extended this policy by introducing
  the \texttt{DEAL\_II\_\allowbreak DEPRECATED\_EARLY} macro, which indicates that a feature will be
  deprecated in the next release. In contrast to the first macro, it will only give
  warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
  \item After each update of the master branch of \dealii{}, we build a new Docker image
  with all features enabled. It is
  accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
  particular useful when used in the contiguous-integration processes of user codes.
\end{itemize}
%
The changelog lists more than 200 other features and bugfixes.

@@

\todo[inline]{TODO}

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.3: updated 5/08/2021 PM
%      sorted by WB 5/28
Pasquale Africa,
Tyler Anderson,
Mathias Anselmann,
Arpit Babbar,
Maximilian Bergbauer,
Nicolas Barnafi,
Bruno Blais,
Michele Bucelli,
Marcus Calhoun-Lopez,
Fabian Castelli, %He provided major contributions to candi for v9.3
Praveen Chandrashekar,
Conrad Clevenger,
David F. Castellanos,
Andrew Davis,
Elias Dejene,
Niklas Fehn,
Menno Fraters,
Ivan Fumagalli,
Daniel Garcia-Sanchez,
Rene Gassmoeller,
Nicola Giuliani,
Krishnakumar Gopalakrishnan,
Alexander Grayver,
Olivier Guevremont,
Jake Harmon,
Graham Harper,
Katharina Kormann,
Christoph Kammer,
Wenyu Lei,
Zhou Lei,
Phillip Mobley,
Nils Much,
Pratik Nayak,
Toni El Geitani Nehme,
Justin O'Connor,
Daniel Paukner,
Sebastian Proell,
Lei Qiao,
Ce Qin,
Reza Rastak,
Julian Roth,
Laura Prieto Saavedra,
Doug Shi-Dong,
David Schneider,
Magdalena Schreter,
Richard Schussnig,
Konrad Simon,
Dominic Soldner,
Simon Sticko,
Malhar Tidke,
Ignacio Tomas,
Peter Westerbaan,
Jiaqi Zhang.

>>>>

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Marc Fehling,
    Rene Gassm{\"o}ller
    Timo Heister,
    Luca Heltai,
    Uwe K{\"o}cher,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Sebastian Proell,
    Konrad Simon,
    Bruno Turcksin,
    David Wells,
    Jiaqi Zhang
  },
  pdftitle={The deal.II Library, Version 9.3, 2021},
}

@@

\author[4]{Rene~Gassm{\"o}ller}
\affil[4]{Department of Geological Sciences,
   University of Florida,
   1843 Stadium Road,
   Gainesville, FL, 32611, USA.
  {\texttt{rene.gassmoeller@ufl.edu}}}

\author[5]{Timo~Heister}
 \affil[5]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister/jiaqi2@clemson.edu}}}

\author[6]{Luca~Heltai}
\affil[6]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{luca.heltai@sissa.it}}}

\author[7]{Uwe~K{\"o}cher}
\affil[7]{Chair of Numerical Mathematics,
  Helmut-Schmidt-University,
  University of the Federal Armed Forces Hamburg,
  Holstenhofweg~85, 22043 Hamburg, Germany.
  {\texttt{uwe.koecher@hsu-hh.de}}}

 \author[8,9]{Martin~Kronbichler}
 \affil[8]{Institute for Computational Mechanics,
   Technical University of Munich,
   Boltzmannstr.~15, 85748 Garching, Germany.
   {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
 \affil[9]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler@it.uu.se}}}

\author[10]{Matthias~Maier}
\affil[10]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[8,11]{Peter Munch}
 \affil[11]{Institute of Materials Research, Materials Mechanics,
 Helmholtz-Zentrum Geesthacht,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hzg.de}}}

\author[12]{Jean-Paul~Pelteret}
\affil[12]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

\author[8]{Sebastian~Proell}

\author[17]{Konrad Simon}
  \affil[17]{Department of Mathematics /
  Center for Earth System Research and Sustainability (CEN),
  University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}

@@

\author[5]{Jiaqi~Zhang}

@@

\dealii{} version 9.3.0 was released June ??, 2021.
\todo{Update date}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.3. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%
In addition, we discuss the \texttt{candi} installation program in Section~\ref{subsec:candi}.

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii{} release
that we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item Each non-artificial cell now has a globally unique index that can be queried
  for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
  via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
  can be used to efficiently index into global vectors storing data
  for each cell, rather than for each degree of freedom corresponding to a finite
  element field.

  \item Previously, functions and classes were marked using the 
  macro \texttt{DEAL\_II\_DEPRECATED} and then typically removed in
  the release after the one in which these deprecation notices were
  available to users. We have now extended this policy by introducing
  the \texttt{DEAL\_II\_\allowbreak DEPRECATED\_EARLY} macro, which indicates that a feature will be
  deprecated in the next release. In contrast to the first macro, it will only give
  warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
  \item After each update of the master branch of \dealii{}, we build a new Docker image
  with all features enabled. It is
  accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
  particular useful when used in the continuous-integration processes of user codes.
\end{itemize}
%
The changelog lists more than 200 other features and bugfixes.

@@

\dealii 's particle infrastructure has been modernized to store nearly all
particle data as a collection of continuous data arrays instead of a container of
objects. This reorganization improves cache efficiency when iterating over
particles and reduces the amount of data that needs to be moved when a particle
moves to a different cell. Keeping track of unused data slots allows to reuse
them for new particles, significantly reducing memory allocations when
particles are created after other particles have left the local domain. The
data arrays are rebuilt and sorted during every mesh refinement cycle.

Additionally, the particle infrastructure now supports a faster search algorithm
for particles that moved farther than one cell width between particle sorting
operations, and support for updating ghost particles (particles that live in
ghost cells around the local domain) by updating their properties instead of
destroying and rebuilding their container. The latter step improves the
efficiency of ghost particle exchange significantly.

Finally, the \texttt{Particles::DataOut} class now supports writing particle
properties as vectors or tensors instead of a collection of scalars if the properties are
marked as such.

@@

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people
contributed code to this release:\\
%
% Uwe Koecher doesn't usually show up in the changelog, but
% we should make sure he's listed.
%
% 9.3: updated 5/08/2021 PM
%      sorted by WB 5/28
Pasquale Africa,
Tyler Anderson,
Mathias Anselmann,
Arpit Babbar,
Maximilian Bergbauer,
Nicolas Barnafi,
Bruno Blais,
Michele Bucelli,
Marcus Calhoun-Lopez,
Fabian Castelli, %He provided major contributions to candi for v9.3
Praveen Chandrashekar,
Conrad Clevenger,
David F. Castellanos,
Andrew Davis,
Elias Dejene,
Niklas Fehn,
Menno Fraters,
Ivan Fumagalli,
Daniel Garcia-Sanchez,
Nicola Giuliani,
Krishnakumar Gopalakrishnan,
Alexander Grayver,
Olivier Guevremont,
Jake Harmon,
Graham Harper,
Katharina Kormann,
Christoph Kammer,
Wenyu Lei,
Zhou Lei,
Phillip Mobley,
Nils Much,
Pratik Nayak,
Toni El Geitani Nehme,
Justin O'Connor,
Daniel Paukner,
Lei Qiao,
Ce Qin,
Reza Rastak,
Julian Roth,
Laura Prieto Saavedra,
Doug Shi-Dong,
David Schneider,
Magdalena Schreter,
Richard Schussnig,
Dominic Soldner,
Simon Sticko,
Malhar Tidke,
Ignacio Tomas,
Peter Westerbaan.
